---
title: "Gender Differences in Chess"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                      out.width = "100%")

library(magrittr)
library(data.table)
library(ggplot2)
library(foreach)
library(dplyr)
library(tidyr)
library(patchwork)
library(doParallel)
registerDoParallel(cores=4)
library(kableExtra)
library(cowplot)
library(ggtext)

theme_set(
  theme_minimal() +
    theme(axis.line = element_line(colour = "grey80"),
          axis.ticks = element_line(colour = "grey80"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_blank(),
          panel.background = element_blank())
)


  

run_permutations <- FALSE
n_draws = 100000
```

# Define a few functions

```{r permutation-functions, eval = run_permutations}
# define functions to generate permutation tests to analyse whether the distribution
# of top women and men can be explained by pure chance

permute <- function(scores_female, scores_male, n_highest = 10, 
                    fct = mean) {
  n_small <- length(scores_female)
  n_large <- length(scores_male)
  n_total <- n_small + n_large
  
  all_scores <- c(scores_female, scores_male)
  
  # shuffle scores
  all_scores <- all_scores[sample(n_total)]
  
  # draw small and large sample
  values_small <- all_scores[1:n_small]
  values_large <- all_scores[(n_small + 1):n_total]
  
  if (n_highest == 1) {
    mean_small <- max(values_small)
    mean_large <- max(values_large)
  } else if (n_highest == Inf) {
    mean_small <- do.call(fct, list(values_small))
    mean_large <- do.call(fct, list(values_large))
  } else {
    mean_small <- do.call(fct, list(tail(sort(values_small), n_highest)))
    mean_large <- do.call(fct, list(tail(sort(values_large), n_highest)))
  }
  return(mean_large - mean_small)
}

apply_permutation <- function(region, data, fct = mean) {
  
  if (region != "all") {
    data <- data %>%
      filter(country == region)
  }
  
  scores_female <- data %>%
    dplyr::filter(sex == "F") %>%
    dplyr::pull(rating)
  
  scores_male <- data %>%
    dplyr::filter(sex == "M") %>%
    dplyr::pull(rating)
  
  observed_difference <- data %>%
    dplyr::group_by(sex) %>%
    dplyr::arrange(rating) %>%
    dplyr::slice_tail(n = n_highest) %>%
    dplyr::summarise(out = do.call(fct, list(rating)), 
                     .groups = "drop_last") %>%
    dplyr::pull(out) %>%
    diff()
  
  permuted_differences <- replicate(n = n_draws, expr = {
    permute(scores_female, scores_male, n_highest = n_highest, fct = fct)
  })
  
  res <- data.frame(country = region, 
                    observed_difference = observed_difference, 
                    mean_perm_difference = mean(permuted_differences), 
                    pval = 1 - mean(permuted_differences <= observed_difference))
  
  return(res)
}


# function to create a histogram for males and females
create_histogram <- function(data) {
  hist_F <- data %>%
    dplyr::filter(sex == "F") %>%
    dplyr::mutate(sex = "Female") %>%
    ggplot(aes(x = rating)) +
    geom_histogram(aes(y = stat(count) / sum(count)), 
                   colour = "white") + 
    facet_wrap(~ sex) + 
    # # theme_minimal() +
    scale_y_continuous(limits = c(0, 0.1)) + 
    scale_x_continuous(limits = c(1000, 2900)) + 
    labs(y = "Proportion", x = "Rating")
  
  hist_M <- data %>%
    dplyr::filter(sex == "M") %>%
    dplyr::mutate(sex = "Male") %>%
    ggplot(aes(x = rating)) +
    geom_histogram(aes(y = stat(count) / sum(count)), 
                   colour = "white") + 
    facet_wrap(~ sex) + 
    # # theme_minimal() + 
    scale_y_continuous(limits = c(0, 0.1)) +
    scale_x_continuous(limits = c(1000, 2900)) +
    labs(y = "Proportion", x = "Rating")
  
  hist_all <- hist_F + hist_M
  
  return(hist_all)
}

scatter_plots <- function(data, top_x = Inf, 
                          pvals_mean_rating, 
                          pvals_median_rating, 
                          pvals_sd_rating) {
  
  get_plot_data <- function(data, summary_fct = mean, pvals) {
    plot_data <- data %>%
      dplyr::group_by(country, sex) %>%
      dplyr::arrange(country, sex, -rating) %>%
      slice(1:min(top_x, nrow(data))) %>%
      dplyr::summarise(rating = summary_fct(rating)) %>%
      tidyr::pivot_wider(names_from = sex,
                         values_from = rating) %>%
      dplyr::mutate(difference = M - F) %>%
      dplyr::inner_join(pvals) %>%
      dplyr::mutate(`*p*-value` = ifelse(p_val < 0.05 | p_val > 0.95,
                                       "Significant",
                                       "Not significant"))
  }
  
  get_lims <- function(plot_data, factor = 100) {
    ylims <- c(floor(min(plot_data$F) / factor) * factor, 
             ceiling(max(plot_data$F) / factor) * factor) 
    xlims <- c(floor(min(plot_data$M) / factor) * factor, 
               ceiling(max(plot_data$M) / factor) * factor) 
    
    lims <- c(min(ylims[1], xlims[1]), 
              max(ylims[2], xlims[2]))
    return(lims)
  }
  
  scatter <- function(plot_data) {
    out <- plot_data %>%
      ggplot(aes(y = F, x = M)) +
      geom_point(aes(colour = `*p*-value`), size = 0.8) +
      geom_abline(aes(slope = 1, intercept = 0), 
                  linetype = "dashed", color = "grey40") +
      coord_cartesian(xlim = lims, ylim = lims) + 
      theme(legend.position = "none") + 
      scale_color_manual(values = c("black", "tomato3")) 
    return(out)
  }
  
  inset_plot <- function(plot_data) {
    plot_data %>%
      ggplot(aes(x = difference)) +
      geom_histogram(aes(y = stat(count) / sum(count)), 
                     colour = "white", fill = "grey50") + 
      geom_vline(xintercept = 0, linetype = "dashed") + 
      scale_y_continuous(limits = c(0, 0.12), breaks = c(0, 0.03, 0.06, 0.09))
  }
  
  plot_data_mean <- get_plot_data(data, 
                                  summary_fct = mean, 
                                  pvals = pvals_mean_rating)
  plot_data_median <- get_plot_data(data, 
                                    summary_fct = median, 
                                    pvals = pvals_median_rating)
  
  lims <- get_lims(plot_data_median)
  
  p1 <- scatter(plot_data_mean) + 
    labs(y = "Mean rating females", x = "Mean rating males")
  
  inset1 <- inset_plot(plot_data_mean) +
    labs(x = "\u0394 mean rating", y = "Proportion")  
  
  # scatter plot for the median of women and men in every country
  # ------------------------------------------------------------------------------
  p2 <- plot_data_median %>%
    scatter() + 
    labs(y = "Median rating females", x = "Median rating males")
  
  # additional histogram
  inset2 <- plot_data_median %>%
    inset_plot() + 
    labs(x = "\u0394 median rating", y = "Proportion")  
  
  # scatter plot for the sd of women and mean in every country
  # ------------------------------------------------------------------------------
  plot_data_sd <- get_plot_data(data, 
                                summary_fct = sd, 
                                pvals = pvals_sd_rating)
  
  lims <- get_lims(plot_data_sd, factor = 10)
  
  p3 <- plot_data_sd %>%
    scatter() + 
    labs(y = "SD rating females", x = "SD rating males") 
  # + 
  #   scale_x_continuous(breaks = c(50, ))
  
  # additional histogram
  inset3 <- plot_data_sd %>%
    inset_plot() + 
    labs(x = "\u0394 SD rating", y = "Proportion")  
  
  if (top_x == 1) {
    p3 <- NULL
    inset3 <- NULL
  }
  
  # together ---------------------------------------------------------------------
  
  layout <- "
  ACE
  BDF
  "
  
  (p1 + inset1 + p2 + inset2 + p3 + inset3) +
    plot_layout(guides = "collect", 
                heights = c(2, 1),
                design = layout) +
    plot_annotation(tag_levels = "A") + 
    theme(legend.position = 'bottom')  + 
    theme(legend.title = element_markdown())
}



# function to create a summary table
summarise_data <- function(data, n_highest = Inf) {
  data %>%
    group_by(country, sex) %>%
    arrange(country, rating) %>%
    slice_tail(n = n_highest) %>%
    group_by(country) %>%
    mutate(n_country = n()) %>%
    group_by(country, sex) %>%
    summarise(mean = mean(rating), 
              median = median(rating),
              sd = sd(rating), 
              median = median(rating),
              n_country = unique(n_country),
              n = n(),
              n_rel = n() / unique(n_country)) %>%
    ungroup() %>%
    tidyr::pivot_wider(id_cols = c(country,n_country), 
                       names_from = sex, 
                       values_from = c(mean, median, sd, n_rel, n)) %>%
    mutate(diff_mean = mean_M - mean_F, 
           diff_median = median_M - median_F)
}
```



# Filter data

```{r}
# data decisions: 
# use fide data list from Dec 31 2019
# data downloaded from fide website on October 1 2021 https://ratings.fide.com/download_lists.phtml 
# include all junior players for main analysis --> no filter on birthday
# exlude every player who has an "i" or an "wi" flag

# read in data
chess <- fread("data/ratings20191231_downloaded20211001.csv")

# compute age from birthday
# filter out
# - those who do not have a sex given
# - those who are inactive
# - countries with fewer than 10 FIDE rated players
worldwide <- chess %>%
  mutate(age = year(Sys.Date()) - birthday) %>%
  group_by(country) %>%
  filter(sex != "", 
         !(flag %in% c("i", "wi"))) 

countries_worldwide <- worldwide$country %>% 
  unique()

  
# for country-by-country data filter countries with less than 30 female players
# technically this is at least either 30 men or women
countries_with_30_females <- worldwide %>%
  filter(sex == "F") %>%
  group_by(country) %>%
  summarise(n = n()) %>%
  filter(n >= 30) %>%
  pull(country)

# World Bank version of these country names
wb_countrycodes <- countrycode::countrycode(
  countries_with_30_females, origin = "ioc", 
  destination = "wb"
) 
wb_countrycodes[is.na(wb_countrycodes)] <- "GBR"
  

country_by_country <- worldwide %>%
  filter(country %in% countries_with_30_females)
```

# Analyis on the worldwide level

```{r}
# numer of countries
worldwide$country %>% 
  unique() %>%
  length()

# number of females and males
worldwide %>%
  group_by(sex) %>%
  summarise(number = n())

# mean ratings men and women
worldwide %>%
  group_by(sex) %>%
  summarise(mean_rating = mean(rating, na.rm = TRUE)) 

# median ratings men and women
worldwide %>%
  group_by(sex) %>%
  summarise(median_rating = median(rating, na.rm = TRUE)) 

# sd ratings
worldwide %>%
  group_by(sex) %>%
  summarise(sd_rating = sd(rating, na.rm = TRUE)) |>
  pull(sd_rating)%>%
  diff()


worldwide |>
  summarise_data()

# histogram of men and women 
hist <- create_histogram(worldwide) + 
  theme(axis.text.y = element_blank(), 
        axis.title.y = element_blank(), 
        axis.line.y = element_blank())

ggsave(plot = hist, file = "results/plots/histograms_ratings_world.png", 
       width = 3.7, height = 2)

# permutation test to compare the means of males and females
if (run_permutations) {
  # done
  n_highest = Inf
  # perm_mean_world <- apply_permutation(region = "all", data = worldwide)
# data.table::fwrite(perm_mean_world, "results/results_mean_world_all.csv")
  
  # not done
  perm_median_world <- apply_permutation(region = "all", data = worldwide, fct = median)
  data.table::fwrite(perm_median_world, "results/results_median_world.csv")
  
  perm_sd_world <- apply_permutation(region = "all", data = worldwide, fct = sd)
  data.table::fwrite(perm_sd_world, "results/results_sd_world.csv")
}

# WMW test on full data
wilcox.test(rating ~ sex, data = worldwide)
```

Top 1 gap for worldwide
```{r}
if (run_permutations) {
  # not done
  n_highest = 1
  perm_mean_world_1 <- apply_permutation(region = "all", data = worldwide)
  data.table::fwrite(perm_mean_world_1, "results/results_mean_world_1.csv")

  # done
  perm_median_world <- apply_permutation(region = "all", data = worldwide, fct = median)
  data.table::fwrite(perm_median_world, "results/results_median_world_1.csv")
  
  perm_sd_world <- apply_permutation(region = "all", data = worldwide, fct = sd)
  data.table::fwrite(perm_sd_world, "results/results_sd_world_1.csv")
}

```


Top 10 gap for worldwide
```{r}
if (run_permutations) {
  # not done
  n_highest = 10
  perm_mean_world_10 <- apply_permutation(region = "all", data = worldwide)
  data.table::fwrite(perm_mean_world_1, "results/results_mean_world_10.csv")

  # not done
  # perm_median_world <- apply_permutation(region = "all", data = worldwide, fct = median)
  # data.table::fwrite(perm_median_world, "results/results_median_world.csv")
  
  # perm_sd_world <- apply_permutation(region = "all", data = worldwide, fct = sd)
  # data.table::fwrite(perm_sd_world, "results/results_sd_world.csv")
}

```


```{r}
# WMW test for the different countries
wmw <- data.table::copy(country_by_country)
setDT(wmw)
wmw <- split(wmw, by = "country")

res <- purrr::map(wmw, .f = function(data) {
  wilcox.test(rating ~ sex, data = data)$p.value
}) |> 
  cbind() |>
  as.data.frame()

res$country <- rownames(res)

res |>
  filter(V1 > 0.05)

# Kolmogorov-Smirnoff-Test on the p-values
ks.test(x = unlist(res$V1), y = "punif")

```

# Permutation analysis for ALL players (using the country_by_country data)

```{r}
# number of countries
country_by_country$country %>% 
  unique() %>%
  length()

# number of countries where mean male was higher than mean female
country_by_country %>%
  group_by(country, sex) %>%
  summarise(mean_rating = mean(rating)) %>%
  pivot_wider(names_from = sex, values_from = mean_rating) %>%
  mutate(male_advantage = M > F) %>%
  pull(male_advantage) %>%
  sum()

# number of countries where median male was higher than mean female
country_by_country %>%
  group_by(country, sex) %>%
  summarise(median_rating = median(rating)) %>%
  pivot_wider(names_from = sex, values_from = median_rating) %>%
  mutate(male_advantage = M > F) %>%
  filter(!male_advantage) 

# number of countries where median male was higher than mean female
country_by_country %>%
  group_by(country, sex) %>%
  summarise(sd_rating = sd(rating)) %>%
  pivot_wider(names_from = sex, values_from = sd_rating) %>%
  mutate(male_higher = M > F) %>%
  filter(male_higher) 


# permutation tests on the mean 
if (run_permutations) {
  # done
  n_highest = Inf
  countries <- unique(country_by_country$country)
  results_all <- foreach(i = 1:length(countries), .combine = 'rbind') %dopar% {
    apply_permutation(region = countries[i], data = country_by_country)
  }
  names(results_all) <- c("country", "Obs. all", "Expected all", "p all")
  data.table::fwrite(results_all, "results/results_all.csv")
}

# permutations median
if (run_permutations) {
  # done
  n_highest = Inf
  countries <- unique(country_by_country$country)
  results_all <- foreach(i = 1:length(countries), .combine = 'rbind') %dopar% {
    apply_permutation(region = countries[i], data = country_by_country, fct = median)
  }
  names(results_all) <- c("country", "Obs. all", "Expected all", "p all")
  data.table::fwrite(results_all, "results/results_all_median.csv")
}

# permutations sd
if (run_permutations) {
  # done
  n_highest = Inf
  countries <- unique(country_by_country$country)
  results_all <- foreach(i = 1:length(countries), .combine = 'rbind') %dopar% {
    apply_permutation(region = countries[i], data = country_by_country, fct = sd)
  }
  names(results_all) <- c("country", "Obs. all", "Expected all", "p all")
  data.table::fwrite(results_all, "results/results_all_sd.csv")
}

# p-values
pvals_mean <- fread("results/results_all.csv") |>
  rename(p_val = `p all`) |>
  select(p_val, country)

# number significant
pvals_mean %>%
  filter(p_val > 0.025)

# Kolmogorov-Smirnoff-test on p-values for the mean
ks.test(x = pvals_mean$p_val, y = "punif")

pvals_median <- fread("results/results_all_median.csv") |>
  rename(p_val = `p all`) |>
  select(p_val, country)

pvals_median %>%
  filter(p_val < 0.025)

# Kolmogorov-Smirnoff-test on p-values for the median
ks.test(x = pvals_median$p_val, y = "punif")

pvals_sd <- fread("results/results_all_sd.csv") |>
  rename(p_val = `p all`) |>
  select(p_val, country)

pvals_sd %>%
  filter(p_val < 0.025 | p_val > 0.975)

# Kolmogorov-Smirnoff-test on p-values for the standard deviation
ks.test(x = pvals_sd$p_val, y = "punif")



# scatter plots with differences and significance
scatter_plots(country_by_country, pvals_mean_rating = pvals_mean, 
              pvals_sd_rating = pvals_sd, pvals_median_rating = pvals_median)

ggsave("results/plots/all_scatter_plots-all.png", height = 4, width = 7)


```


# Permutation analyses top 10 players using the country by country data

```{r}
# permutation tests on the mean 
if (run_permutations) {
  n_highest = 10
  countries <- unique(country_by_country$country)
  results_top10 <- foreach(i = 1:length(countries), .combine = 'rbind') %dopar% {
    apply_permutation(region = countries[i], data = country_by_country)
  }
  names(results_top10) <- c("country", "Obs. all", "Expected all", "p all")
  data.table::fwrite(results_top10, "results/results_top10_mean.csv")
}

# permutations median
if (run_permutations) {
  n_highest = 10
  countries <- unique(country_by_country$country)
  results_all <- foreach(i = 1:length(countries), .combine = 'rbind') %dopar% {
    apply_permutation(region = countries[i], data = country_by_country, fct = median)
  }
  names(results_all) <- c("country", "Obs. all", "Expected all", "p all")
  data.table::fwrite(results_all, "results/results_top10_median.csv")
}

# permutations sd
if (run_permutations) {
  n_highest = 10
  countries <- unique(country_by_country$country)
  results_all <- foreach(i = 1:length(countries), .combine = 'rbind') %dopar% {
    apply_permutation(region = countries[i], data = country_by_country, fct = sd)
  }
  names(results_all) <- c("country", "Obs. all", "Expected all", "p all")
  data.table::fwrite(results_all, "results/results_top10_sd.csv")
}


results_all <- fread("results/results_top10_mean.csv")
results_all %>%
  filter(`p all` < 0.025 | `p all` > 0.975) %>%
  pull(`p all`) %>%
  length()


pvals_mean <- fread("results/results_top10_mean.csv") |>
  rename(p_val = `p all`) |>
  select(p_val, country)

pvals_median <- fread("results/results_top10_median.csv") |>
  rename(p_val = `p all`) |>
  select(p_val, country)

pvals_sd <- fread("results/results_top10_sd.csv") |>
  rename(p_val = `p all`) |>
  select(p_val, country)

# scatter plots with differences and significance
scatter_plots(country_by_country, pvals_mean_rating = pvals_mean, 
              pvals_sd_rating = pvals_sd, pvals_median_rating = pvals_median, 
              top_x = 10)

ggsave("results/plots/all_scatter_plots-top10.png", height = 4, width = 7)
```



# Permutation analyses top 1 players using the country by country data

```{r}
# permutation tests on the mean 
if (run_permutations) {
  n_highest = 1
  countries <- unique(country_by_country$country)
  results_top1 <- foreach(i = 1:length(countries), .combine = 'rbind') %dopar% {
    apply_permutation(region = countries[i], data = country_by_country)
  }
  names(results_top1) <- c("country", "Obs. all", "Expected all", "p all")
  data.table::fwrite(results_top1, "results/results_top1_mean.csv")
}

# permutations median
if (run_permutations) {
  n_highest = 1
  countries <- unique(country_by_country$country)
  results_top1 <- foreach(i = 1:length(countries), .combine = 'rbind') %dopar% {
    apply_permutation(region = countries[i], data = country_by_country, fct = median)
  }
  names(results_top1) <- c("country", "Obs. all", "Expected all", "p all")
  data.table::fwrite(results_top1, "results/results_top1_median.csv")
}

# permutations sd
if (run_permutations) {
  n_highest = 1
  countries <- unique(country_by_country$country)
  results_top1 <- foreach(i = 1:length(countries), .combine = 'rbind') %dopar% {
    apply_permutation(region = countries[i], data = country_by_country, fct = sd)
  }
  names(results_top1) <- c("country", "Obs. all", "Expected all", "p all")
  data.table::fwrite(results_top1, "results/results_top1_sd.csv")
}


results_all <- fread("results/results_top1_mean.csv")
# number of significant countries
results_all %>%
  filter(`p all` < 0.025 | `p all` > 0.975)

# mean difference expected vs. real
results_all |>
  summarise(diff = mean(`Expected all` - `Obs. all`), 
            diff_sd = sd(`Expected all` - `Obs. all`))

pvals_mean <- fread("results/results_top1_mean.csv") |>
  rename(p_val = `p all`) |>
  select(p_val, country)

ks.test(pvals_mean$p_val, y = "punif")

pvals_median <- fread("results/results_top1_median.csv") |>
  rename(p_val = `p all`) |>
  select(p_val, country)

pvals_sd <- fread("results/results_top1_sd.csv") |>
  rename(p_val = `p all`) |>
  select(p_val, country)

# scatter plots with differences and significance
scatter_plots(country_by_country, pvals_mean_rating = pvals_mean, 
              pvals_sd_rating = pvals_sd, pvals_median_rating = pvals_median, 
              top_x = 1)

ggsave("results/plots/all_scatter_plots-top1.png", height = 4, width = 7)


```




# Difference in mean performance vs. share of female players



```{r}
# scatter plot of average performance difference vs. female participation 
country_by_country %>%
  summarise_data() %>%
  ggplot(aes(x = diff_mean, y = n_rel_F)) + 
  geom_point(size = 0.8) + 
  labs(x = "\u0394 mean rating M - F", 
       y = "Percentage of female players in a country") + 
  geom_smooth(method = "glm", 
              method.args=list(family="binomial"), 
              alpha = 0.1, colour = 'grey', size = 0.4) + 
  # theme_minimal() + 
  coord_flip() + 
  scale_y_continuous(labels = function(x) paste0(x*100, "%")) + 
  expand_limits(x = 0) + 
  geom_vline(xintercept = 0, linetype = "dashed")

ggsave("results/plots/female-participation-performance.png", 
       width = 3.5, height = 2)

# check correlation between difference in performance and share of women
df <- country_by_country %>%
  summarise_data() %>%
  mutate(n = n_F + n_M) %>%
  select(diff_mean, diff_median, n_rel_F, n) 

# mean
cor.test(df$diff_mean, df$n_rel_F)
# run a weighted least squares regression 
model <- lm(diff_mean ~ n_rel_F, data = df, weights = n)
summary(model)

# median correlation
cor.test(df$diff_median, df$n_rel_F)

# same for median
country_by_country %>%
  summarise_data() %>%
  ggplot(aes(x = diff_median, y = n_rel_F)) + 
  geom_point() + 
  labs(x = "\u0394 median rating men - women", 
       y = "Percentage of female players in a country") + 
  geom_smooth(method = "glm", 
              method.args=list(family="binomial"), 
              alpha = 0.1, colour = 'grey', size = 0.4) + 
  # theme_minimal() + 
  coord_flip() + 
  scale_y_continuous(labels = function(x) paste0(x*100, "%")) + 
  expand_limits(x = 0) + 
  geom_vline(xintercept = 0, linetype = "dashed")


ggsave("results/plots/female-participation-performance-median.png", 
       width = 7, height = 4)

```


# Permutation tests






```{r}
# mean age plot for the countries ----------------------------------------------
age_men_women <- country_by_country %>%
  dplyr::group_by(country, sex) %>%
  dplyr::arrange(-rating) %>%
  dplyr::slice_head(n = 10) %>%
  dplyr::summarise(age = mean(2021 - birthday, na.rm = TRUE)) %>%
  tidyr::pivot_wider(names_from = sex, 
                     values_from = age) %>%
  dplyr::mutate(difference = M - F) %>%
  ggplot(aes(y = F, x = M)) +
  geom_point(size = 0.8) +
  geom_abline(aes(slope = 1, intercept = 0), 
              linetype = "dashed", color = "grey40") +
  coord_cartesian(xlim = c(18, 47), ylim = c(18, 47)) + 
  # theme_minimal() + 
  labs(y = "Mean age female", x = "Mean age male")


# relationship between age and rating ------------------------------------------
# for top 10: age vs. ratings
agetop10 <- country_by_country %>%
  dplyr::group_by(country, sex) %>%
  dplyr::arrange(-rating) %>%
  dplyr::slice_head(n = 10) %>%
  dplyr::mutate(age = 2021 - birthday) %>%
  summarise(age = mean(age), rating = mean(rating)) %>%
  rename(Sex = sex) %>%
  ggplot(aes(y = rating, x = age, group = Sex, fill = Sex, color = Sex)) + 
  geom_smooth(alpha = 0.1, size = 0.5) + 
  geom_point(size = 0.8) + 
  scale_color_manual(values = c("olivedrab", "royalblue3")) + 
  scale_fill_manual(values = c("olivedrab", "royalblue3")) + 
  labs(y = "Rating top 10", x = "Mean age top 10")

ggsave("results/plots/scatter_age_rating.png")

# plot with age difference vs. rating difference for top 10. not so sure
rating_vs_age_top10 <- country_by_country %>%
  dplyr::group_by(country, sex) %>%
  dplyr::arrange(-rating) %>%
  dplyr::slice_head(n = 10) %>%
  dplyr::mutate(age = 2021 - birthday) %>%
  dplyr::summarise(age = mean(age, na.rm = TRUE), 
                   rating = mean(rating)) %>%
  tidyr::pivot_wider(names_from = sex, values_from = c(age, rating)) %>%
  dplyr::mutate(age_diff = age_M - age_F, 
                rating_diff = rating_M - rating_F, 
                `Avg. man older` = age_diff > 0) %>%
  ggplot(aes(y = rating_diff, x = age_diff)) + 
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") + 
  geom_point(size = 0.8) + 
  # geom_point(aes(color = `Avg. man older`)) + 
  geom_smooth(method = "lm", se = TRUE, 
              size = 0.4,
              colour = 'grey', alpha = 0.1) + 
  scale_x_continuous(breaks = c(-5, 0, 5, 10, 15)) + 
  labs(y = "\u0394 mean rating M - F", x = "\u0394 mean age M - F")



# combine age plots
age_men_women + agetop10 +  rating_vs_age_top10 +
  plot_layout(widths = c(1, 1, 1.0), guides = "collect") &
  theme(legend.position = "bottom") &
  plot_annotation(tag_levels = 'A')

ggsave("results/plots/scatter_combined_age.png", 
       width = 7, height = 3)
```
































# Summary tables

```{r}
make_summary_table <- function(data_summary) {
  data_summary %>%
  mutate(n_rel_F = n_rel_F * 100) %>%
  select(Country = country, 
         N = n_country, 
         "N \\textit{M}" = n_M, 
         `N \\textit{F}` = n_F,
         `\\% \\textit{F}` = n_rel_F,
         `Mean \\textit{M}` = mean_M, 
         `Mean \\textit{F}` = mean_F, 
         `sd \\textit{M}` = sd_M,
         `sd \\textit{F}` = sd_F
         ) %>%
  mutate_if(is.numeric, round, digits = 1) %>%
  kable(format = "latex", escape = FALSE) %>%
  kable_styling()
}
```

Summary table of all players

```{r}
data_summary_all <- summarise_data(data)
make_summary_table(data_summary_all)
```


Summary table of top 10 players

```{r}
data_summary_10 <- summarise_data(data, n_highest = 10)
make_summary_table(data_summary_10)
```




