---
title: "Gender Differences in Chess"
output: github_document
---

## Background

Professor Wei Ji Ma recently posted an [article](https://en.chessbase.com/post/what-gender-gap-in-chess) on [chessbase](https://chessbase.com) in which he analyzed gender differences in the performance of chess players in India. 

His basic argument is this: Often, people claim that women are worse at chess and point to the low number of female chess players among the all the top chess players in the world. While the observation is correct, according to Professor Ma, the conclusion is false. The reason is that comparing the maximum values is a very bad way of comparing two distributions. He finds that indeed women in India are likely overperforming given what we should expect from the low overall number of women in chess. 

In the following I replicate his analysis and also look at the entire world, not only India. A blog post on the topic can be found (here)[https://followtheargument.org/gender-differences-among-top-performers-in-chess]. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(magrittr)
library(data.table)
library(ggplot2)
chess <- readr::read_csv("chess_standard_dataframe.csv")

# set parameters
n_highest = 10
n_draws = 100000
redo_from_scratch <- FALSE

# filter datasets
chess_filtered_world <- chess %>%
  dplyr::filter(birthday <= 2000)

chess_filtered_india <- chess %>%
  dplyr::filter(birthday <= 2000, 
                country == "IND") 

```

## Replication: Analysis of gender differences in India

### Relative performance of men and women in India

```{r india}
plot <- ggplot()

plot <- plot + 
  geom_density(data = chess_filtered_india %>%
                 dplyr::filter(sex == "F"), 
               aes(x = rating, y = ..count../sum(..count..), color = sex), alpha = 1) +   
  geom_density(data = chess_filtered_india %>%
                 dplyr::filter(sex == "M"), 
               aes(x = rating, y = ..count../sum(..count..), color = sex), alpha = 1) + 
  labs(y = "probability", x = "Elo Rating") + 
  theme_light()

plot 
```

### Basic summary statistics India
```{r summary-india}
summary_india <- chess_filtered_india %>%
  dplyr::group_by(sex) %>%
  dplyr::summarise(number = max(dplyr::n()), 
                   mean = mean(rating), 
                   median = median(rating), 
                   sd = sd(rating)) %>%
  dplyr::ungroup()
summary_india
```

### Performance of best players in India
```{r top-india}
summary_india_top <- chess_filtered_india %>%
  dplyr::group_by(sex) %>%
  dplyr::arrange(sex, rating) %>%
  dplyr::slice_tail(n = n_highest) %>%
  dplyr::summarise(number = max(dplyr::n()),
                   mean = mean(rating), 
                   median = median(rating), 
                   sd = sd(rating)) %>%
  dplyr::ungroup()

summary_india_top
```

```{r preparation-india}
# store values needed later in the analysis
n_small_india <- summary_india %>%
  dplyr::filter(sex == "F") %>%
  dplyr::pull(number)

n_large_india <- summary_india %>%
  dplyr::filter(sex == "M") %>%
  dplyr::pull(number)

difference_top_india <- summary_india_top %>%
  dplyr::pull(mean) %>%
  diff()

# define permutation function 
permute <- function(df, n_small = 20012, n_large = 243375, n_highest = 1) {
  n_total <- nrow(df)
  
  # resort data.frame
  df <- df[sample(n_total), ]
  
  values_small <- df$rating[1:n_small]
  values_large <- df$rating[(n_small + 1):n_total]
  
  if (n_highest == 1) {
    mean_small <- max(values_small)
    mean_large <- max(values_large)
  } else {
    mean_small <- mean(tail(sort(values_small), n_highest))
    mean_large <- mean(tail(sort(values_large), n_highest))
  }
  return(data.frame(mean_small, mean_large))
}

```

```{r permutation-analysis-india}

if (redo_from_scratch) {
  l_india <- list()
  #convert to data.table for speed
  df <- data.table::as.data.table(chess_filtered_india)[, .(rating)] 
  for (i in 1:n_draws) {
    l_india[[i]] <- permute(df, 
                            n_small = n_small_india, 
                            n_large = n_large_india, 
                            n_highest = n_highest)
  }
  
  results_india <- dplyr::bind_rows(l_india) %>%
    dplyr::mutate(difference = mean_large - mean_small)
  
  data.table::fwrite(results_india, paste0("permutation_india_top", n_highest, ".csv"))

} else {
  results_india <- readr::read_csv(paste0("permutation_india_top", n_highest, ".csv"))
}
  
```

### Actual difference top-performers
```{r actual-difference-india}
difference_top_india
```

### Mean difference top-performers in permutation analysis
```{r mean-difference-india}
results_india %>%
  dplyr::pull(difference) %>%
  mean()
```

### Distribution of differences in permuataion analysis
```{r distribution-india}

percent_smaller <- results_india %>%
  dplyr::filter(difference <= difference_top_india) %>%
  nrow() / nrow(results_india)

caption <- paste0("The observed difference is larger than ", 
                  round(percent_smaller * 100, 1), " percent of simulated differences")

results_india %>%
  ggplot() + 
  geom_density(aes(x = difference, y = ..count..)) + 
  geom_vline(xintercept = difference_top_india, color = "red") + 
  labs(caption = caption) + 
  theme_light() 

```





  
  
## New Analysis: Gender Differences in the World

### Relative performance of men and women in the world

```{r world}
plot <- ggplot()

plot <- plot + 
  geom_density(data = chess_filtered_world %>%
                 dplyr::filter(sex == "F"), 
               aes(x = rating, y = ..count../sum(..count..), color = sex), alpha = 1) +   
  geom_density(data = chess_filtered_world %>%
                 dplyr::filter(sex == "M"), 
               aes(x = rating, y = ..count../sum(..count..), color = sex), alpha = 1) + 
  labs(y = "probability", x = "Elo Rating") + 
  theme_light()

plot 
```

### Basic summary statistics
```{r summary-world}
summary_world <- chess_filtered_world %>%
  dplyr::group_by(sex) %>%
  dplyr::summarise(number = max(dplyr::n()), 
                   mean = mean(rating), 
                   median = median(rating), 
                   sd = sd(rating)) %>%
  dplyr::ungroup()
summary_world
```

### Performance of top players in the world
```{r top-world}
summary_world_top <- chess_filtered_world %>%
  dplyr::group_by(sex) %>%
  dplyr::arrange(sex, rating) %>%
  dplyr::slice_tail(n = n_highest) %>%
  dplyr::summarise(number = max(dplyr::n()), 
                   mean = mean(rating), 
                   median = median(rating), 
                   sd = sd(rating)) %>%
  dplyr::ungroup()
summary_world_top
```

```{r preparation-world}
# store values needed later in the analysis
n_small_world <- summary_world %>%
  dplyr::filter(sex == "F") %>%
  dplyr::pull(number)

n_large_world <- summary_world %>%
  dplyr::filter(sex == "M") %>%
  dplyr::pull(number)

difference_top_world <- summary_world_top %>%
  dplyr::pull(mean) %>%
  diff()

```

```{r permutation-analysis-world}

if (redo_from_scratch) {
  l_world <- list()
  #convert to data.table for speed
  df <- data.table::as.data.table(chess_filtered_world)[, .(rating)] 
  for (i in 1:n_draws) {
    l_world[[i]] <- permute(df, 
                            n_small = n_small_world, 
                            n_large = n_large_world, 
                            n_highest = n_highest)
  }
  
  results_world <- dplyr::bind_rows(l_world) %>%
    dplyr::mutate(difference = mean_large - mean_small)
  
  data.table::fwrite(results_world, paste0("permutation_world_top", n_highest, ".csv"))

} else {
  results_world <- readr::read_csv(paste0("permutation_world_top", n_highest, ".csv"))
}
  
```

### Actual difference top-performers
```{r actual-difference-world}
difference_top_world
```

### Mean difference top-performers in permutation analysis
```{r mean-difference-world}
results_world %>%
  dplyr::pull(difference) %>%
  mean()
```

### Distribution of differences in permuataion analysis
```{r distribution-world}

percent_smaller <- results_world %>%
  dplyr::filter(difference <= difference_top_world) %>%
  nrow() / nrow(results_world)

caption <- paste0("The observed difference is larger than ", 
                  round(percent_smaller * 100, 1), " percent of simulated differences")

results_world %>%
  ggplot() + 
  geom_density(aes(x = difference, y = ..count..)) + 
  geom_vline(xintercept = difference_top_world, color = "red") + 
  labs(caption = caption) + 
  theme_light()

```

## Differences in median performances across countries

Only countries with 1000 or more registered players are taken into account
``` {r country-performance}

chess_filtered_world %>%
  dplyr::group_by(country) %>%
  dplyr::filter(dplyr::n() > 3000) %>%
  dplyr::group_by(country, sex) %>%
  dplyr::summarise(median = median(rating)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(country) %>%
  dplyr::summarise(difference = diff(median)) %>%
  dplyr::arrange(difference) %>%
  dplyr::mutate(country = forcats::fct_reorder(country, sort(difference))) %>%
  ggplot(aes(y = difference, x = country, group = 1)) + 
  geom_point() + 
  geom_line(alpha = 0.3) + 
  theme_minimal()

```

### Correlation between performance differences and the proportion of female players

```{r correlation}

median_difference <- chess_filtered_world %>%
  dplyr::group_by(country, sex) %>%
  dplyr::summarise(median = median(rating)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(country) %>%
  dplyr::summarise(difference = diff(median))

percentage_women <- chess_filtered_world %>%
  dplyr::group_by(country) %>%
  dplyr::summarise(women = mean(sex == "F")) 

combined <- dplyr::full_join(median_difference, percentage_women) %>%
  dplyr::filter(!is.na(difference))

cor(combined$difference, combined$women)

```


``` {r world-map, eval = FALSE}
library(highcharter)
library(countrycode)

plot_map <- combined %>%
  dplyr::mutate(country_iso = countrycode(country, 
                                      origin = "ioc", 
                                      destination = "iso3c"), 
                country = ifelse(is.na(country_iso), country, country_iso), 
                country = ifelse(country == "ENG", "GBR", country)) %>%
  dplyr::rename("iso-a3" = country)


map <- hcmap(
  map = "custom/world-highres3", # high resolution world map
  data = plot_map, # name of dataset
  joinBy = "iso-a3",
  value = "difference",
  showInLegend = TRUE, # hide legend
  nullColor = "#DADADA",
  download_map_data = TRUE
) %>%
  hc_mapNavigation(enabled = TRUE) %>%
  hc_legend("none") %>%
  hc_title(text = "World map")

library(htmlwidgets)
library(highcharter)

saveWidget(map, file = "map.html")
```